Push-Pull Signal-Function Functional Reactive Programming

### 2012-11-29 Docs/IFL2012/design.tex
edits to 2.0

The goal of FRP is to provide an efficient, declarative abstraction for creating
reactive programs. Towards this overall goal, there are three goals which this
system is intended to meet.

@@

@@

It is fine for a system to be composable with regards to itself, but an FRP
system must interact with the outside world. Since we cannot anticipate every
possible form of input and output that the system will be asked to interact
with, we must interface with Haskell's IO system. In particular, most libraries
for user interaction (e.g. GUI and graphics libraries such as GTK+ and GLUT) and
most libraries for time-dependent IO (e.g. audio and video systems) make use of
the event loop abstraction. In this abstraction, event handlers are registered
with the system, and then a command is issued to run a loop which detects events
and runs the handlers, and uses the results of the handlers to render the
appropriate output.

We would like for the FRP system to be easy to integrate with such IO systems,
while being flexible enough to enable its use with other forms of IO systems,
such as simple imperative systems, threaded systems, or network servers.

>>>>

The purpose of FRP is to provide an efficient, declarative abstraction for creating
reactive programs. To this end, there are three goals which the TimeFlies
system is intended to meet. \rn{What are the three!?  Enumerate.  Push
based, composable, IO-system agnostic?}

@@

\todo{At this point we still need a clear notion of what a pull-based
  evaluation of events would actually do.}

@@

\rn{Can the above be made into a slightly more technical claim?
  Composable abstractions enable composing $O(N)$ components while
  reasoning about the $O(N)$ local interactions, rather than the
  $O(N^2)$ global interactions between all components?  Second, I
  think not needing to know the implementation is separate from
  ``composability''; surely there is some name for that: ADTs,
  ``encapsulation'', information hiding, etc.}

% It is fine for a system to be composable with regards to itself, but
% an FRP system must
\new{Further, an FRP program is not a closed system; it must}
interact with the outside world. Since we cannot anticipate every
possible form of input and output that the system will be asked to interact
with, we must interface with Haskell's IO system. 
\rn{First mention of Haskell... need some transition?}
In particular, most libraries
for user interaction (e.g. GUI and graphics libraries such as GTK+ and GLUT) and
most libraries for time-dependent IO (e.g. audio and video systems) make use of
the event-loop abstraction. In this abstraction, event handlers are registered
\new{centrally}; 
\new{once initiated, the running event-loop detects events and invokes handlers; and
  handlers are responsible for rendering output.}
%% and then a command is issued to run a loop which detects events
%% and runs the handlers, and uses the results of the handlers to render the
%% appropriate output.

\rn{Is the event loop really an {\em abstraction}?  (What is
  abstracted?)  Or is it just a ``model'' or an API?}

We would like for the FRP system to be easy to integrate with such IO systems,
while being flexible enough to enable its use with other forms of IO systems,
such as simple imperative systems, threaded systems, or network servers.
\rn{vague... not sure what these mean.  Threaded = multiple event
  loop?  Alternative to an event loop is what... blocking IO operations?}
